---
title: "Meta-analysis Criterion validity"
author: "Stefano Berrone"
date: "2025-09-10"
output:
  pdf_document: default
  html_document: default
---

## Introduction

This is part of a meta-analysis of a larger systematic review about the measurement properties of the Borg scales. In this analysis, we meta-analysed the correlation coefficients (treated as effect size) of the studies that about *Criterion Validity* of the *Borg 6-20 scale*.

## Code & output


### Install packages

```{r, install/load packages }
install.packages(c("readxl", "dplyr", "metafor"), dependencies = TRUE)

library(readxl)
library(dplyr)
library(metafor)

```


### Import data

```{r, Import data}
borg <- read_excel("Data_Borg.xlsx")

library(knitr)
kable(borg, caption = "Borg dataset")
```


### Cleaning (ensure numeric r and n)

```{r, Data cleaning}
to_num <- function(x) {
  if (is.numeric(x)) return(x)
  x <- gsub(",", ".", as.character(x))
  suppressWarnings(as.numeric(x))
}

borg <- borg %>%
  mutate(
    r            = to_num(r),
    n            = to_num(n),
    validity_type = as.character(validity_type),
    comparator    = as.character(comparator),
    prom          = as.character(prom),
    author_year   = as.character(author_year)
  )

```

### Analysis. Filter: criterion validity + Borg 6–20

```{r}
criterion <- borg %>%
  filter(validity_type == "criterion") %>%
  filter(prom != "BorgCR10")
```


### Compute Fisher's Z effect sizes for r

```{r}
criterion_es <- escalc(
  measure = "ZCOR",
  ri = r,
  ni = n,
  data = criterion
)
```


### Fit random-effects meta-analysis

```{r}
res <- rma(yi, vi, data = criterion_es, method = "REML")

print(summary(res))
print(predict(res, transf = transf.ztor))
```


### Forest plot 

```{r}
slabs <- criterion$author_year
forest(
  res,
  slab   = slabs,
  xlab   = "Correlation (r)",
  transf = transf.ztor
)

```




















